---
title: "ALT Sports Analysis"
author: "Callan Capitolo"
output:
  html_document:
    self_contained: false
    toc: true
    toc_float: true
    number_sections: false
    theme:
      version: 4
      bootswatch: flatly
    code_folding: hide
    code_download: true  
css: styles.css
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,   # hide all code
  message    = FALSE,   # suppress package messages
  warning    = FALSE,   # suppress warnings
  fig.width  = 6,       # nice figure size
  fig.height = 4,
  dpi        = 300,     # high-res output
  fig.pos    = "H",     # place plots here
  out.width  = "100%"   # full-width figures
)
```

```{r}

library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)
library(gt)
library(gtExtras)
library(scales)
race_info <- read.csv("race_info_2024.csv") #overall race info
race_sessions <- read.csv("race_sessions_2024.csv") #information about each session in race
session_competitors <- read.csv("session_competitors_2024.csv") #competitor results
start_list <- read.csv("start_list.csv")

# keeps the first occurrence of each ID, with all other columns
race_info_unique <- race_info %>% 
  distinct()

race_sessions_unique <- race_sessions %>% 
  distinct() %>% 
  rename("Sessions_Name" = Name)

race_competitors_unique <- session_competitors %>% 
  distinct()

clean_racing_df <- race_competitors_unique %>% 
  rename("race_competitors_ID" = ID) %>% 
  left_join(race_sessions_unique %>% rename("race_sessions_ID" = ID) , by = c("SessionID" = "race_sessions_ID", "RaceID")) %>% 
  left_join(race_info_unique %>%  rename("race_info_ID" = ID), by = c("RaceID" = "race_info_ID")) %>% 
  mutate(Position = as.numeric(Position))
```
# Part 1

# Question 1
```{r}
regulars <- clean_racing_df %>% 
  filter(season == 2024) %>% 
  mutate(total_races = n_distinct(RaceID)) %>% 
  group_by(fullName) %>% 
  summarize(races = n_distinct(RaceID), total_races = max(total_races), participation_rate = races/total_races) %>% 
  filter(participation_rate >= 0.75) %>% 
  mutate(regular = "Yes")

regulars %>% 
  select(fullName,participation_rate) %>% 
  arrange(-participation_rate) %>% 
  gt() %>% 
  cols_align(align = "center") %>%
  cols_label(fullName = "Driver",participation_rate = "Participation Rate") %>% 
  gt_theme_538() %>% 
  tab_header(
    title = md("Regular Drivers"),
    subtitle = md("Must Have Race Participation Rate of At Least 75%")
  ) %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )
```

\newpage
# Question 2
```{r}
regular_racing_df <- clean_racing_df %>% 
  filter(season == 2024) %>% 
  left_join(regulars %>% select(fullName,regular), by = c("fullName"))

race_results <- regular_racing_df %>% 
  filter(regular == "Yes") %>% 
  filter(str_detect(Sessions_Name,"A Feature")) %>% #check this
  group_by(fullName) %>% 
  summarize(wins = mean(Position == 1), top_3 = mean(Position <=3), top_5 = mean(Position <= 5), top_10 = mean(Position <= 10))

race_results %>% 
  gt() %>% 
  fmt_number(
    columns = c(wins, top_3, top_5, top_10),
    decimals = 2
  ) %>% 
  cols_align(align = "center") %>%
  cols_label(fullName = "Driver",wins = "Win Rate", top_3 = "Top 3 Rate", top_5 = "Top 5 Rate", top_10 = "Top 10 Rate") %>% 
  gt_theme_538() %>% 
  data_color(
    columns  = -fullName,
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 1)
    )) %>% 
  tab_header(
    title = md("Regular Drivers Race Results")
  ) %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )
```

# Question 3
```{r}
h2h_all <- regular_racing_df %>%
  filter(regular == "Yes") %>% 
  filter(str_detect(Sessions_Name,"A Feature")) %>%
  select(RaceID,SessionID,fullName,Position) %>% 
  inner_join(regular_racing_df %>%   filter(regular == "Yes") %>% 
               filter(str_detect(Sessions_Name,"A Feature")) %>% select(RaceID,SessionID,fullName,Position), by = c("RaceID","SessionID"), suffix = (c(".A",".B"))) %>% 
  mutate(
    A_beats_B = case_when(
      Position.A < Position.B              ~ 1,
      Position.A > Position.B              ~ 0,
      Position.A == Position.B             ~ NA_real_,  # tie
      TRUE                                  ~ NA_real_
    )
  )

h2h_matrix <- h2h_all %>%
  group_by(fullName.A, fullName.B) %>%
  summarise(
    races_together = n(),
    wins           = sum(A_beats_B, na.rm = TRUE),
    losses         = sum(1 - A_beats_B, na.rm = TRUE),
    win_pct        = wins / races_together,
    .groups        = "drop"
  )


# 1. Pivot to a wide matrix of win_pct
h2h_wide <- h2h_matrix %>% 
  select(fullName.A, fullName.B, win_pct) %>% 
  pivot_wider(
    names_from  = fullName.B,
    values_from = win_pct
  ) %>% 
  mutate(
    across(
      -fullName.A,        
      ~ na_if(.x, 0)
    )
  )

# 2. Feed to gt and style
h2h_wide %>% 
  gt(rowname_col = "fullName.A") %>% 
  fmt_percent(
    columns = everything(),     # all opponent columns
    decimals = 0
  ) %>% 
  data_color(
    columns  = everything(),
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 1)
    )
  ) %>% 
  tab_header(
    title    = "Head-to-Head Win%",
    subtitle = "Row beat column this % of the time"
  ) %>% 
  cols_align(
    align = "center",
    columns = everything()
  ) %>% 
  gt_theme_538() %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )
```
AI assisted in help to brainstorming idea for matrix.

# Question 4
```{r}
non_regular_results <- regular_racing_df %>% 
  filter(is.na(regular)) %>% 
  filter(str_detect(Sessions_Name,"A Feature")) %>% #check this
  group_by(fullName) %>% 
  summarize(wins = mean(Position == 1), top_3 = mean(Position <=3), top_5 = mean(Position <= 5), top_10 = mean(Position <= 10), 
            races_count = n_distinct(RaceID)) %>% 
  filter(fullName != "")

non_regular_results %>% 
  filter(races_count >=5) %>% 
  gt() %>% 
  fmt_number(
    columns = c(wins, top_3, top_5, top_10),
    decimals = 2
  ) %>% 
  cols_align(align = "center") %>%
  cols_label(fullName = "Driver", races_count = "Races" ,wins = "Win Rate", top_3 = "Top 3 Rate", top_5 = "Top 5 Rate", top_10 = "Top 10 Rate") %>% 
  gt_theme_538() %>% 
  data_color(
    columns  = c(-fullName,-races_count),
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 1)
    )) %>% 
  data_color(
    columns  = c(races_count),
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(5,20)
    )) %>% 
  tab_header(
    title = md("Non-Regular Drivers Race Results"),
    subtitle = md("Must Haved Competed At Least 5 Races")
  ) %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )

```

# Question 5
```{r}
all_results <- regular_racing_df %>% 
  filter(str_detect(Sessions_Name,"A Feature")) %>% #check this
  group_by(fullName) %>% 
  summarize(races_count = n_distinct(RaceID) ,wins = mean(Position == 1), top_3 = mean(Position <=3), top_5 = mean(Position <= 5), top_10 = mean(Position <= 10)
            ) %>% 
  filter(fullName != "")

start_list_results <- start_list %>% 
  left_join(all_results, by = "fullName") %>% 
  mutate(composite_metric = wins*10+top_3*7+top_5*5+top_10)

start_list_results %>% 
  arrange(-composite_metric) %>% 
  select(-isRegular) %>% 
  filter(races_count >5) %>% 
  slice_head(n=10) %>% 
  gt() %>% 
  fmt_number(
    columns = c(wins, top_3, top_5, top_10, composite_metric),
    decimals = 2
  ) %>% 
  cols_align(align = "center") %>%
  cols_label(fullName = "Driver", races_count = "Races" ,wins = "Win Rate", top_3 = "Top 3 Rate", top_5 = "Top 5 Rate", top_10 = "Top 10 Rate",
             composite_metric = "Composite Metric") %>% 
  gt_theme_538() %>% 
  data_color(
    columns  = c(-fullName,-races_count),
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 1)
    )) %>% 
  tab_header(
    title = md("Event Top Contenders"),
    subtitle = md("Must Haved Competed At Least 5 Races")
  ) %>% 
  data_color(
    columns  = composite_metric,
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 10)
    )) %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )
```
I created a composite metric to create an all-in-one metric to evaluate drivers. The composite metric rewards drivers for wins, but also for consistency through finishing in the top 3, top 5, etc. The formula is **composite metric = win rate * 10 + top 3 rate * 7 +top 5 rate * 5 + top 10**. Through the composite metric I was able to determine 3 tiers of contenders. **Top Contenders**: Brad Sweet, Kyle Larson, and Tyler Courtney - **Middle Contenders**: Rico Abreu, Corey Day - **Low Contenders**: Justin Sanders, Brent Marks, Aaorn Reutzel, Tanner Thorson, Justin Peck

# Question 6
```{r}
regular25 <- start_list %>% filter(isRegular == 1) %>% 
  left_join(regulars, by = "fullName") %>% 
  filter(is.na(regular)) %>% 
  left_join(all_results, by = "fullName")

notes_df <- tribble(
  ~fullName,         ~notes,
  "Aaron Reutzel",   "Three-time All Star Circuit of Champions champ (2018–20); 2 wins in rookie 2021 World of Outlaws season",
  "Chase Randall",   "2021 USAC National Midget ROY & 360 Knoxville champ; 6 wins in 10 Sprint Invaders starts",
  "Daison Pursley",  "2017 USAC Micro Sprint champ; Passing Master in 2023 & 2024; first ever to win Midget, Sprint & Silver Crown in one year (2024)",
  "Danny Sams III",  "Kart & quarter-midget grad; 3 wins in 2024 Ohio Sprint Speedweek + a 360 win at Lawrenceburg",
  "Sye Lynch",       "FAST on Dirt feature winner (2019, 2022); runner-up in Commonwealth Clash; made 2021 Outlaws debut"
)

regular25 %>% select(fullName,races_count,wins,top_3,top_5,top_10) %>% 
  left_join(notes_df, by = "fullName") %>% 
  gt() %>% 
  fmt_number(
    columns = c(wins, top_3, top_5, top_10),
    decimals = 2
  ) %>% 
  cols_align(align = "center") %>%
  cols_label(fullName = "Driver", races_count = "Races" ,wins = "Win Rate", top_3 = "Top 3 Rate", top_5 = "Top 5 Rate", top_10 = "Top 10 Rate", notes = "AI Other Experience Notes") %>% 
  gt_theme_538() %>% 
  data_color(
    columns  = c(-fullName,-races_count, -notes),
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 1)
    )) %>% 
  tab_header(
    title = md("New Regular Drivers Race Results"),
    subtitle = md("Must Haved Competed At Least 5 Races")
  ) %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )
```
I have included some statistics above for the new 2025 regulars and an AI generated summary of their other racing experience.

# Part 2
```{r}
f1_2024 <- read.csv("f1_2024.csv") %>% mutate(position = as.numeric(position)) %>%mutate(driverName = ifelse(driverName == "Alexander Albon", "Alex Albon", 
                                                                                                             ifelse(driverName == "Guanyu Zhou", "Zhou Guanyu",driverName)))#results
f1_2025 <- read.csv("f1_2025.csv") %>% mutate(position = as.numeric(position)) #results
f1_odds_2024 <- read.csv("f1_odds_24.csv") 
```

# Question 1
```{r, fig.width=18,fig.height=10, out.width='100%', fig.align='center', fig.show='hold'}
clean_odds <- f1_odds_2024 %>% mutate(probability = 1/EU.Odds) %>% group_by(Race,Market) %>% mutate(total_market_probability = sum(probability)) %>% ungroup() %>% 
  mutate(no_vig_prob = ifelse(Market == "Top 6", probability * (6/total_market_probability),
                              ifelse(Market == "Top 10",probability * (10/total_market_probability),
                              ifelse(Market == "Event Podium", probability * (3/total_market_probability),
                                     probability/total_market_probability)))) %>% 
  mutate(
    Name.Result = case_when(
      # Map any Red Bull variants to the official "Red Bull Racing"
      Name.Result %in% c("Red Bull")  ~ "Red Bull Racing",
      # Map old Sauber label to the current team name
      Name.Result == "Sauber"  ~ "Kick Sauber",
      # Collapse any Williams variant into simply "Williams"
      Name.Result == "Williams 201"~ "Williams",
      Name.Result == "Visa Cash App RB"~"RB",
      # Otherwise keep the name as-is (e.g. Ferrari, Mercedes, etc.)
      TRUE~ Name.Result
    )
  )


results_vs_prob <- clean_odds %>%   filter(Market %in% c("Event Winner","Event Podium","Top 6","Top 10")) %>%  filter(Name.Result != "") %>% 
  left_join(f1_2024 %>% filter(sessionName == "Race") %>% select(season, eventName,eventNum,driverName,position) %>% 
              group_by(driverName) %>% mutate(races_completed = sum(position >0,na.rm = T)) %>%  ungroup(), 
            by = c("Season" = "season", "RaceStop" = "eventNum", "Name.Result" = "driverName")) %>% 
  mutate(result = ifelse(Market == "Event Winner" & position == 1, 1, ifelse(Market == "Event Podium" & position <= 3, 1, ifelse(Market == "Top 6" & position <= 6, 1,
                                                                                                                                ifelse(Market == "Top 10" & position <= 10, 1,0))))) %>% 
  filter(!is.na(result),!is.na(no_vig_prob)) %>% 
  arrange(RaceStop) %>% 
  group_by(Name.Result,Market) %>% 
  mutate(total_prob = cumsum(no_vig_prob), actual_result = cumsum(result), races = max(races_completed, na.rm = T)) %>% 
  filter(races>=18)

racer <- "Lewis Hamilton"

results_vs_prob %>% pivot_longer(c(total_prob,actual_result), names_to = "Type", values_to = "Values") %>% 
  filter(Name.Result == racer) %>% 
  ggplot(aes(x = RaceStop, y = Values, color = Market, linetype = Type))+
  geom_line(lwd = 4)+
  labs(x = "Race Number", y = "Actual Finishes vs Expected Finishes", title = paste0(racer," Actual Finishes vs Expected Finishes by Race"), 
       caption = "Lines de-vigged through proportional scaling approach")+
  theme(legend.position = "top",
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "white", color="white"),
        legend.title = element_blank(),
        legend.text = element_text(colour = "black", face = "bold", size = 20),
        plot.title = element_text(hjust = .5, colour = "white", face = "bold", size = 26),
        plot.subtitle = element_text(hjust = .5, colour = "white", size = 22),
        plot.caption = element_text(colour = "white", size = 22),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "black", color="black"),
        panel.background = element_rect(fill = "black", color="black"),
        axis.ticks = element_line(color = "white"),
        axis.text = element_text(face = "bold", colour = "white",size = 22),
        axis.title = element_text(color = "white", size = 24),
        panel.border = element_rect(colour = "white", fill = NA, size = 1))+
  scale_linetype_manual(name   = "Series",
                        values = c(total_prob   = "dotted",
                                   actual_result = "solid"),
                        labels = c(total_prob = "Implied No Vig Probability",
                                   actual_result = "Cumulative Results")) +
  scale_color_manual( values = c("Event Podium"   = "purple",  # cyan
                       "Event Winner"   = "red",  # magenta
                       "Top 10"  = "orange",  # yellow
                       "Top 6"   = "darkgrey"   # lime green
  ))+
  guides(
    linetype = guide_legend(
      override.aes = list(
        size = 10,           # make legend lines thicker
        color = "white",
        lwd = 2
      )),
    color = guide_legend(
      override.aes = list(
        size = 10,           # make legend lines thicker
        lwd = 2
      ))
  )

```

```{r, fig.width=18,fig.height=10, out.width='100%', fig.align='center', fig.show='hold'}
racer <- "George Russell"

results_vs_prob %>% pivot_longer(c(total_prob,actual_result), names_to = "Type", values_to = "Values") %>% 
  filter(Name.Result == racer) %>% 
  ggplot(aes(x = RaceStop, y = Values, color = Market, linetype = Type))+
  geom_line(lwd = 4)+
  labs(x = "Race Number", y = "Actual Finishes vs Expected Finishes", title = paste0(racer," Actual Finishes vs Expected Finishes by Race"), 
       caption = "Lines de-vigged through proportional scaling approach")+
  theme(legend.position = "top",
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "white", color="white"),
        legend.title = element_blank(),
        legend.text = element_text(colour = "black", face = "bold", size = 20),
        plot.title = element_text(hjust = .5, colour = "white", face = "bold", size = 26),
        plot.subtitle = element_text(hjust = .5, colour = "white", size = 22),
        plot.caption = element_text(colour = "white", size = 22),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "black", color="black"),
        panel.background = element_rect(fill = "black", color="black"),
        axis.ticks = element_line(color = "white"),
        axis.text = element_text(face = "bold", colour = "white",size = 22),
        axis.title = element_text(color = "white", size = 24),
        panel.border = element_rect(colour = "white", fill = NA, size = 1))+
  scale_linetype_manual(name   = "Series",
                        values = c(total_prob   = "dotted",
                                   actual_result = "solid"),
                        labels = c(total_prob = "Implied No Vig Probability",
                                   actual_result = "Cumulative Results")) +
  scale_color_manual( values = c("Event Podium"   = "purple",  # cyan
                       "Event Winner"   = "red",  # magenta
                       "Top 10"  = "orange",  # yellow
                       "Top 6"   = "darkgrey"   # lime green
  ))+
  guides(
    linetype = guide_legend(
      override.aes = list(
        size = 10,           # make legend lines thicker
        color = "white",
        lwd = 2
      )),
    color = guide_legend(
      override.aes = list(
        size = 10,           # make legend lines thicker
        lwd = 2
      ))
  )
```



```{r, fig.width=18,fig.height=10, out.width='100%', fig.align='center', fig.show='hold'}
racer <- "Sergio Perez"

results_vs_prob %>% pivot_longer(c(total_prob,actual_result), names_to = "Type", values_to = "Values") %>% 
  filter(Name.Result == racer) %>% 
  ggplot(aes(x = RaceStop, y = Values, color = Market, linetype = Type))+
  geom_line(lwd = 4)+
  labs(x = "Race Number", y = "Actual Finishes vs Expected Finishes", title = paste0(racer," Actual Finishes vs Expected Finishes by Race"), 
       caption = "Lines de-vigged through proportional scaling approach")+
  theme(legend.position = "top",
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "white", color="white"),
        legend.title = element_blank(),
        legend.text = element_text(colour = "black", face = "bold", size = 20),
        plot.title = element_text(hjust = .5, colour = "white", face = "bold", size = 26),
        plot.subtitle = element_text(hjust = .5, colour = "white", size = 22),
        plot.caption = element_text(colour = "white", size = 22),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "black", color="black"),
        panel.background = element_rect(fill = "black", color="black"),
        axis.ticks = element_line(color = "white"),
        axis.text = element_text(face = "bold", colour = "white",size = 22),
        axis.title = element_text(color = "white", size = 24),
        panel.border = element_rect(colour = "white", fill = NA, size = 1))+
  scale_linetype_manual(name   = "Series",
                        values = c(total_prob   = "dotted",
                                   actual_result = "solid"),
                        labels = c(total_prob = "Implied No Vig Probability",
                                   actual_result = "Cumulative Results")) +
  scale_color_manual( values = c("Event Podium"   = "purple",  # cyan
                       "Event Winner"   = "red",  # magenta
                       "Top 10"  = "orange",  # yellow
                       "Top 6"   = "darkgrey"   # lime green
  ))+
  guides(
    linetype = guide_legend(
      override.aes = list(
        size = 10,           # make legend lines thicker
        color = "white",
        lwd = 2
      )),
    color = guide_legend(
      override.aes = list(
        size = 10,           # make legend lines thicker
        lwd = 2
      ))
  )
```
I de-vigged the lines through the proportional scaling approach to get a true probability for each market type.

# Question 2
```{r}
max_odds <- clean_odds %>%   filter(Market %in% c("Event Winner","Event Podium","Top 6","Top 10")) %>%  filter(Name.Result != "") %>% 
  left_join(f1_2024 %>% filter(sessionName == "Race") %>% select(season, eventName,eventNum,driverName,position), 
              by = c("Season" = "season", "RaceStop" = "eventNum", "Name.Result" = "driverName")) %>% 
  mutate(result = ifelse(Market == "Event Winner" & position == 1, 1, ifelse(Market == "Event Podium" & position <= 3, 1, ifelse(Market == "Top 6" & position <= 6, 1,
                                                                                                                                 ifelse(Market == "Top 10" & position <= 10, 1,0))))) %>% 
  group_by(Market) %>% 
  summarize(max_odds = max(EU.Odds[result == 1],na.rm = T))

highest_scoring_results <- f1_2024 %>% 
  filter(sessionType == "Race") %>% 
  mutate(
    f1_pts = case_when(
      position == 1  ~ 25,
      position == 2  ~ 18,
      position == 3  ~ 15,
      position == 4  ~ 12,
      position == 5  ~ 10,
      position == 6  ~ 8,
      position == 7  ~ 6,
      position == 8  ~ 4,
      position == 9  ~ 2,
      position == 10 ~ 1,
      TRUE           ~ 0
    )
  ) %>% 
  group_by(eventNum,constructorName) %>% 
  summarize(points = sum(f1_pts,na.rm = T), .groups = "drop") %>% 
  group_by(eventNum) %>% 
  mutate(
    is_top_constructor = points == max(points)
  ) %>% 
  ungroup()

highest_scoring_team <- clean_odds %>% filter(Market == "Highest Scoring Team") %>% 
  left_join(highest_scoring_results, by = c("RaceStop" = "eventNum","Name.Result" = "constructorName")) %>% 
  group_by(Market) %>% 
  summarize(max_odds = max(EU.Odds[is_top_constructor == TRUE],na.rm = T))

margin_of_victory <- f1_2024 %>% filter(sessionType == "Race", position == 2) %>%  group_by(eventNum) %>% summarize(margin = min(gapToLeader,na.rm = T)) %>% 
  mutate(classified_time = case_when(
    margin < 5 ~ "Under 5 seconds",
    margin >=5 & margin <=10 ~ "Between 5-10 seconds",
    TRUE ~ "Over 10 seconds"
  ))

max_margin_odds <- clean_odds %>% filter(Market == "Winning Margin") %>% inner_join(margin_of_victory, by = c("RaceStop" = "eventNum", "Name.Result" = "classified_time")) %>% 
  group_by(Market) %>% 
  summarize(max_odds = max(EU.Odds,na.rm = T))
rbind(max_odds,highest_scoring_team,max_margin_odds) %>% 
  gt() %>% 
  cols_align(align = "center") %>%
  cols_label(max_odds = "Max Odds") %>% 
  gt_theme_538() %>% 
  tab_header(title = "Longest Successful Odds in 2024")
```
# Question 3
```{r}
results_25 <- f1_2025 %>% 
  filter(sessionName == "Race") %>% 
  group_by(driverName) %>% 
  summarize(races = n_distinct(eventKey),wins = mean(position == 1,na.rm =T), top_3 = mean(position <=3,na.rm =T), 
            top_5 = mean(position <= 5,na.rm =T), top_10 = mean(position <= 10,na.rm =T)) %>% 
  mutate(composite_metric = wins*10+top_3*7+top_5*5+top_10)
results_25 %>% 
  arrange(-composite_metric) %>% 
  gt() %>% 
  fmt_number(
    columns = c(wins, top_3, top_5, top_10,composite_metric),
    decimals = 2
  ) %>% 
  cols_align(align = "center") %>%
  cols_label(driverName = "Driver", races = "Races" ,wins = "Win Rate", top_3 = "Top 3 Rate", top_5 = "Top 5 Rate", top_10 = "Top 10 Rate", composite_metric = "Composite Metric") %>% 
  gt_theme_538() %>% 
  data_color(
    columns  = c(-driverName,-races,-composite_metric),
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 1)
    )) %>% 
    data_color(
    columns  = composite_metric,
    # map low->high win_pct to a color ramp
    colors   = col_numeric(
      palette = c("white", "darkgreen"),
      domain  = c(0, 16)
    )) %>% 
  tab_header(
    title = md("F1 2025 Results")
  ) %>% 
  tab_options(
    heading.align = "center"    # centers both title and subtitle
  )
```

I have again used the composite metric defined earlier to analyze these drivers. There is 2 angles I would look to play. The first angle is using the composite metric to identify Charles Leclerc as a driver with value. The reason his composite metric is much lower than the top 4 is because he is yet to have a win this year, but his top 3, top 5, and top 10 results are similar to the top 4. The top 4 are priced between +185 to +450 while Leclerc is priced at +1400 which is a significant jump. This odds change is too big to Leclerc which is why there is value in taking him at +1400. The other angle I would look to play is Max Verstappen due to the recent exit of Christian Horner causing uncertainty in the potential impact of this firing. I do not think books are exactly sure how this change will effect Verstappen's performance and I want to lean into this uncertainty.  With the race being months away, I want to lean into this uncertainty early through betting Verstappen now before the markets can fully understand the firing impact.